---
title: "Class 11"
author: "Emmanuel Robles"
format: pdf
---

# Section 1. Proportion of G|G in a population. 
>   Q1: What are those 4 candidate SNPs?

(rs12936231, rs8067378, rs9303277, and rs7216389)

> Q2: What three genes do these variants overlap or effect?

rs930… IKZF3
rs721… GSDMB
rs806… none

> Q3: What is the location of rs8067378 and what are the different alleles for rs8067378?

Chromosome 17:39895095 (forward strand)

> Q4: Name at least 3 downstream genes for rs8067378?

ORMDL3
GSDMB
GSDMA



Downloaded csv file from Ensamble. Now we'll read the csv file.
```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
```

```{r}
table(mxl$Genotype..forward.strand.) / nrow(mxl) * 100
```

> Q5: What proportion of the Mexican Ancestry in Los Angeles sample population (MXL) are homozygous for the asthma associated SNP (G|G)?

14% of the sample population are homozygous for the asthma SNP.



Now let's look at a different population. GBR.
```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
```

Find proportion of G|G
```{r}
round(table(gbr$Genotype..forward.strand.) / nrow(gbr) *100, 2)
```

>  Q6. Back on the ENSEMBLE page, use the “search for a sample” field above to find the particular sample HG00109. This is a male from the GBR population group. What is the genotype for this sample?

G|G



This variant that is associated with childhood asthma is more frequent in the bgr population than the MXL population. 
Let's dig further.


# Section 2

>  Q7: How many sequences are there in the first file? What is the file size and format of the data? Make sure the format is fastqsanger here!

3,863

>  Q8: What is the GC content and sequence length of the second fastq file?

53%

> Q9: How about per base sequence quality? Does any base have a mean quality score below 20?

All the mean quality scores for the bases are at 30 or higher. No.


# Section 3

> Q10: Where are most the accepted hits located?

Most of his are located around ~38,060,000 and ~38,150,000

> Q11: Following Q10, is there any interesting gene around that area?

For the first we see ORMDL3 an for the one furthest away we see PSMD3


>  Q12: Cufflinks again produces multiple output files that you can inspect from your right-hand-side galaxy history. From the “gene expression” output, what is the FPKM for the ORMDL3 gene? What are the other genes with above zero FPKM values?

128189









