---
title: "Class 11 - Candy"
author: "Emmanuel Robles"
format: pdf
---


```{r}
candy <- read.csv("candy-data.csv", row.names= 1)
head(candy)
```


> Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```


> Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```


> What are these fruity candies?

```{r}
rownames( candy[ candy$fruity == 1, ])
rownames( candy[ candy$chocolate == 1, ])

```


## How often does my favorite candy win?


> Q3. What is your favorite candy in the dataset and what is it’s winpercent value? Mine is higher than professors. Superior candy, clearly.

```{r}
candy["Twix","winpercent"]
candy["Reese's Peanut Butter cup", "winpercent"]
```


> Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat","winpercent"]

```


> Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars","winpercent"]
```


```{r}
library("skimr")
skim(candy)
skimr::skim(candy)
```


> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

Yes, the `winpercent` column is on a 0:100 scale and all others appear to be on a 0:1 scale.


> Q7. What do you think a zero and one represent for the candy$chocolate column?

A zero here means a candy is not classified as containing chocolate.


> Q8. Plot a histogram of winpercent values

```{r}
library(ggplot2)

ggplot(candy, aes(winpercent)) +
  geom_histogram(bins=10)
```


> Q9. Is the distribution of winpercent values symmetrical?

No


> Q10. Is the center of the distribution above or below 50%?

Below 50% with a mean of 
```{r}
mean(candy$winpercent)
```


> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

Fom the averages it looks like people prefer chocolate candy over fruity.
```{r}
fruit.mean <- mean(candy[candy$fruity == 1, ]$winpercent)
choco.mean <- mean(candy[candy$chocolate == 1, ]$winpercent)
fruit.mean
choco.mean
```
 

```{r}
t.test(candy[candy$chocolate == 1, ]$winpercent, candy[candy$fruity == 1, ]$winpercent)
```


## Overall Candy Rankings
There is a base function in R called `sort()` for, guess what, sorting vector inputs.
> Q13. What are the five least liked candy types in this set?

```{r}
candy.rank <- sort(candy$winpercent, decreasing= TRUE)
candy.rank
```


The buddy funciton to `sort()` this is often more useful is called `order()`.

```{r}

```


I can order by `winpercent`.
```{r}
ord <- order(candy$winpercent)
ord
head(candy[ord,])
```


> Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
ord1  <- order(candy$winpercent, decreasing= TRUE)
head(candy[ord1,])
```


> Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)

ggplot(candy) +
  aes(winpercent, rownames(candy)) +
  geom_col()
```


> Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```


```{r}
library(ggplot2)

ggplot(candy) +
  aes(winpercent, reorder(rownames(candy), winpercent)) +
  geom_col(fill=my_cols)

```


> Q17. What is the worst ranked chocolate candy?

Sixlets


> Q18. What is the best ranked fruity candy?

Starburst


## Taking a look at pricepoint


> What is the best candy for the least money?

```{r}
ggplot(candy)+ 
  aes(winpercent, pricepercent)+ 
  geom_point(col=my_cols)
```

```{r}
library(ggrepel)
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```


> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?



## 5 Exploring the correlation structure

Pearson correlation goes between -1 and +1 with zero indicating no correlation and values close to one being very highly (an) correlated


```{r}
library(corrplot)

cij <- cor(candy)
corrplot(cij)
```


> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

Chocolate and fruit.

> Q23. Similarly, what two variables are most positively correlated?

Chocolate and winpercent or chocolate and bar.


## 6 Principal Component Analysis.

The base R function for PCA is called `prcomp()` and we can set "scale= TRUE/FALSE".

```{r}
pca <- prcomp(candy, scale= TRUE)
summary(pca)
```

The main result of PCA - i.e. the new PC plot (projection of candy in our new PC axis) is contained in `pca$x`

```{r}
pc <- as.data.frame(pca$x)

ggplot(pc)+
  aes(PC1, PC2)+
  geom_point(col=my_cols)
```










