---
title: "Class 10"
author: "Emmanuel Robles"
format: html
---

> Q10. Which of the packages above is found only on BioConductor and not CRAN? `msa`

> Q11. Which of the above packages is not found on BioConductor or CRAN?: `bio3d-view`

> Q12. True or False? Functions from the devtools package can be used to install packages from GitHub and BitBucket? `TRUE`




```{r}
library(bio3d)

aa <- get.seq("1ake_A")
aa
```

> Q13. How many amino acids are in this sequence, i.e. how long is this sequence? `214`

We can run a BLAST with this sequence. 

```{r}
#b <- blast.pdb(aa)
```

```{r}
#hits <- plot(b)
```

Let's see what's in our `hits` object

```{r}
#hits$pdb.id

hits <- NULL
hits$pdb.id <- c("1AKE_A", "4X8M_A", "6S36_A", "6RZE_A", "4X8H_A", "3HPR_A", "1E4V_A", "5EJE_A", "1E4Y_A", "3X2S_A", "6HAP_A", "6HAM_A", "4K46_A", "4NP6_A", "3GMT_A", "4PZL_A")
```

Now we can download all these PDB structure files:

```{r}
# Download releated PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

```{r}

```


Now I want to align and superpose these structures which are all over the place.

```{r}
# Align releated PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```

```{r}
pdbs
```

Now we have our aligned superposed structures we can perform all sorts of analysis on them. Let's do PCA...

```{r}
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)
```

Results of PCA on Adenylate kinase X-ray structures. Each dot represents one PDB structure. 

We can cluster the structures by RMSD (or any other method)

```{r}
rd <- rmsd(pdbs)

# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)

```


We can make a movie - also called a trajectory of the major differences (i.e. structural displacements) of ADK 

```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```


























