---
title: "Class 17"
author: "Emmanuel"
format: pdf
---

Pertussis (whooping cough) is a highly contagious lung infection - `B. pertussis`
The CDC has tracked these cases since the 1920's.

```{r}
cdc <- data.frame(
                                 Year = c(1922L,1923L,1924L,1925L,
                                          1926L,1927L,1928L,1929L,1930L,1931L,
                                          1932L,1933L,1934L,1935L,1936L,
                                          1937L,1938L,1939L,1940L,1941L,1942L,
                                          1943L,1944L,1945L,1946L,1947L,
                                          1948L,1949L,1950L,1951L,1952L,
                                          1953L,1954L,1955L,1956L,1957L,1958L,
                                          1959L,1960L,1961L,1962L,1963L,
                                          1964L,1965L,1966L,1967L,1968L,1969L,
                                          1970L,1971L,1972L,1973L,1974L,
                                          1975L,1976L,1977L,1978L,1979L,1980L,
                                          1981L,1982L,1983L,1984L,1985L,
                                          1986L,1987L,1988L,1989L,1990L,
                                          1991L,1992L,1993L,1994L,1995L,1996L,
                                          1997L,1998L,1999L,2000L,2001L,
                                          2002L,2003L,2004L,2005L,2006L,2007L,
                                          2008L,2009L,2010L,2011L,2012L,
                                          2013L,2014L,2015L,2016L,2017L,2018L,
                                          2019L,2020L,2021L),
         Cases = c(107473,164191,165418,152003,
                                          202210,181411,161799,197371,
                                          166914,172559,215343,179135,265269,
                                          180518,147237,214652,227319,103188,
                                          183866,222202,191383,191890,109873,
                                          133792,109860,156517,74715,69479,
                                          120718,68687,45030,37129,60886,
                                          62786,31732,28295,32148,40005,
                                          14809,11468,17749,17135,13005,6799,
                                          7717,9718,4810,3285,4249,3036,
                                          3287,1759,2402,1738,1010,2177,2063,
                                          1623,1730,1248,1895,2463,2276,
                                          3589,4195,2823,3450,4157,4570,
                                          2719,4083,6586,4617,5137,7796,6564,
                                          7405,7298,7867,7580,9771,11647,
                                          25827,25616,15632,10454,13278,
                                          16858,27550,18719,48277,28639,32971,
                                          20762,17972,18975,15609,18617,
                                          6124,2116)
       )
```

Plotting the number of reported cases per year in the U.S.

```{r}
library(ggplot2)
ggplot(cdc)+
  aes(x= Year, y= Cases)+
  geom_point()+
  geom_line()
```

First big whole-cell vaccine program - 1942

```{r}
library(ggplot2)
ggplot(cdc)+
  aes(x= Year, y= Cases)+
  geom_point()+
  geom_line()+
  geom_vline(xintercept= 1942, linetype= 2, color= "red")+
  geom_vline(xintercept= 1980, linetype= 2, color= "grey")+
  geom_vline(xintercept= 1995, linetype= 2, color= "blue")


```

Something is happening with pertusis. Big outbreaks and more cases. Vaccine efficacy is considered as a possible answer. CMI-PB project is studying this issue.

Data is in JSON format ("key:value" pair style)
We will use "jsonlight" package to read data.
```{r}
library(jsonlite)
subject <- read_json("https://www.cmi-pb.org/api/subject", simplifyVector = TRUE) 
head(subject)
```

> Q4. How may aP and wP infancy vaccinated subjects are in the dataset?

```{r}
table(subject$infancy_vac)
```


> Q5. How many Male and Female subjects/patients are in the dataset?

```{r}
table(subject$biological_sex)
```


> Q6. What is the breakdown of race and biological sex (e.g. number of Asian females, White males etc…)?

```{r}
table(subject$race)
```

```{r}
table(subject$race, subject$biological_sex)
```

Reading more data from CMI-PB

```{r}
specimen <- read_json("http://cmi-pb.org/api/specimen", simplifyVector = TRUE)
head(specimen)
```


I want to join the subject and specimen tables together.
```{r}
library(dplyr)

meta <- inner_join(subject, specimen)
head(meta)
```

```{r}
library(jsonlite)

ab <- read_json("http://cmi-pb.org/api/ab_titer", simplifyVector = TRUE)

head(ab)
```

Now I can join "meta" with the "ab". 
```{r}
abdata <- inner_join(meta, ab)
```

```{r}
dim(abdata)
```

> Q11. How many specimens (i.e. entries in abdata) do we have for each isotype?

```{r}
table(abdata$isotype)
```

> Q12. What do you notice about the number of visit 8 specimens compared to other visits?

```{r}
table(abdata$visit)
```

Much less visit 8 specimens because the project is still ongoing and we haven't collected all the patient data.

## Examine IgG1

```{r}
ig1 <- abdata %>% filter(isotype == "IgG1", visit!=8)
head(ig1)

```

```{r}
library(ggplot2)
ggplot(ig1) +
  aes(MFI, antigen) +
  geom_boxplot() + 
  facet_wrap(vars(visit), nrow=2)
```

FIM2/3 changes. Fimbral protein - bacterial pilus and cell adhesion.
















